---
title: "Practical 1"
output:
  html_notebook: 
    highlight: espresso
    css: dlnbstyle.css
    theme: sandstone
    toc: yes
  html_document:
    df_print: paged
    css: "dlnbstyle.css"
    toc: true
editor_options:
  chunk_output_type: inline
---

We start logistic regression, it will help us understand the meaning of non-linearity. 
 

First, few things that we need by way of setup:
```{r, echo= T}
knitr::opts_chunk$set(
  comment = '',
  fig.width = 6,
  fig.height = 6,
  cache = FALSE
)
# for generic R help and tutorials
# help.start()
#  Setup  --------------------
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# install.packages("magrittr")
# install.packages("psych")
library(magrittr)
library(MASS)
library(psych)
```

# Warm up: logistic regression

We use the Boston dataset. Let's look at it
```{r, eval=TRUE, echo=TRUE}
#  Warm up: Logistic regression  ------------
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
names(Boston) %>% print
# Print the first 3 lines
head(Boston,3) %>% print

```

Use `?Boston` for more details on this dataset

Let's split the dependent variable to two parts (say cheap and expensive).


```{r, eval= T}
med_value <- median(Boston$medv)
# predict how many house will have higher price than the median house price
y <- ifelse(Boston$medv > med_value, 1, 0)

# Add out target to the data, and remove the median from the data 
Boston$y <- y
Boston$medv <- NULL

```

Divide data into Train and test

```{r}
set.seed(11)
TT <- NROW(Boston)
tmp_ind <- sample(1:TT, size = TT / 2)

trainn <- Boston[tmp_ind, ]
testt <- Boston[-tmp_ind, ]

dim(trainn); dim(testt);
```

Now let's train logistic regression. 
```{r}
# Logistic regression

logreg <- glm(y ~ ., data = trainn, family = binomial)
summary(logreg)
```
Run the following code to see the effect of the variable `dis` on the probability of the house being expensive.

```{r}
dis_range <- seq(
  from = min(Boston$dis), to = max(Boston$dis),
  by = .1
)
n_range <- length(dis_range)

coeff <- coef(logreg)[-9] # remove 'dis'
mean_val <- apply(Boston, 2, mean)[-c(8, 14)] # without 'dis' and 'y'
mean_val <- c(1, mean_val) # add a constant
y0 <- rep(coeff %*% mean_val, n_range) +
  coef(logreg)[9] * dis_range
prob0 <- exp(y0) / (1 + exp(y0))

plot(dis_range, prob0,
  ylim = c(0, 1), las = 1,
  type = "l", lwd = 3, lty = 1, col = "blue",
  xlab = "Weighted mean of distances to five Boston employment centres", ylab = "",
  main = "Change in Probability"
)
grid(col = "grey", lwd = 2)

```

Let's do the same for the variable `age`:

```{r}
age_range <- seq(from = min(Boston$age), to = max(Boston$age), by = .1)
n_range_age <- length(age_range)

coeff2 <- coef(logreg)[-8] # remove 'age'
mean_val2 <- apply(Boston, 2, mean)[-c(7, 14)] # without 'dis' and 'y'
mean_val2 <- c(1, mean_val2)
y0_2 <- rep(coeff2 %*% mean_val2, n_range_age) + coef(logreg)[8] * age_range

prob0_2 <- exp(y0_2) / (1 + exp(y0_2))

plot(dis_range, prob0,
  ylim = c(0, 1), las = 1, xlim = c(1, 100),
  type = "l", lwd = 3, lty = 1, col = "blue",
  xlab = "Impact of distance and age on the probability", ylab = "",
  main = "Change in Probability"
)
lines(age_range, prob0_2, lwd = 3, lty = 1, col = "green"); grid(col = "grey", lwd = 2)
```
## Predict
```{r}
# get the fit using the model build using train dataset
trainn$fitt <- predict(logreg, data = trainn, type = "response")

# Predict using the model build using test dataset

testt$pred <-
  predict(logreg, newdata = testt, type = "response")
# names(testt)
# confusion matrix
confuse_tab <- table(actual = testt$y, predict = testt$pred > 0.5)
(confuse_tab[1, 1] + confuse_tab[2, 2]) / sum(confuse_tab)
```


# Activation funcions

Code the sigmoid function.

```{r}
# install.packages("Deriv")
library(Deriv)
sigmoid <- function(x) {
  X <- 1 / (1 + exp(-x))
  X
}
sigmoid(0.2)

curve(sigmoid, from = -5, to = 5)

grad_sigmoid <- Deriv(sigmoid)
grad_sigmoid(0.5)

curve(grad_sigmoid, from = -5, to = 5)
```

Code the relu function:
```{r}

relu <- function(x) {
  ifelse(x > 0, x, 0)
}
curve(relu, from = -5, to = 5)

grad_relu <- Deriv(relu)
curve(grad_relu, from = -5, to = 5)

```
Code the swish function

```{r}
swish <- function(x) {
  x * sigmoid(x)
}
curve(swish, from = -5, to = 5)
grad_swish <- Deriv(swish)
curve(grad_swish, from = -10, to = 10)
```



## what does the softmax do?

```{r}
TT <- 1000
x <- runif(TT, -10, 10)
par(mfrow = c(2, 1))
plot(x, pch = 19, ylab = "x")
tmpp <- tail(order(x), 5)
points(x[tmpp] ~ tmpp, pch = 19, las = 1, ylab = "", col = 2)
grid()
plot(exp(x), pch = 19, cex=0.5) # we can divide exp(x) if we want to normalize
grid()
points(exp(x)[tmpp] ~ tmpp, pch = 19, col = 2)

```

